Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
fixing typo in figure

As explained in Section~XXX, we expect to achieve a performance of
about 210Gflop/s in single precision in the best case. We will focus
on the single precision case with 4 vectors and 4 matrices. But
similar techniques apply for other combinations. This target
performance represents 10\% of the peak performance of the
architecture which is impossible without an efficient vectorization of
the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
instruction on fully loaded registers must be executed at most every
10 cycles.

@@

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for the 4 non zero of
the matrix at once. The 4 floats coming from the four matrices for
each non zero are naturally grouped within each lane of the SIMD
vector. One can use a swizzle operation to broadcast the 4 elements of
a single matrix to take the whole vector.

Loading the entries from the {\tt x} vector is a little bit more
complicated. First the columns pointers of 4 different non zeroes are
loaded in a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to unpack for the
  LSB and MSB. Both are mandatory even according the documentation of
  the hardware}. They replicated so that each fill a vector lane using
a swizzle operation. Then each value is multiplied by 4 (because there
are 4 vectors) and the offsets $(3,2,1,0)$ is added each lane to
obtain the correct index of the elements of the {\tt x} vector within
each lane. Then a {\tt gather} operation is performed bring the 16
entries of the {\tt x} vector into the SIMD register. At this point we
have a SIMD register where each lane is the four vector entries for a
non zero element. Using a broadcast permutation, one can replicate an
entry of the four vectors accross each lane.

%\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}

%\begin{figure}[h]
%\begin{center}
%\lstset{language=c++,basicstyle=\scriptsize,numbers=left,numbersep=5pt}
%\begin{lstlisting}{float=h, Caption}
%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
%int int_mask = 0x1111;
%__mmask16 mask = _mm512_int2mask(int_mask);
%
%#pragma omp parallel for
%for (int r=0; r < nb_rows; r++) {
%  __m512 accu = _mm512_setzero_ps();
%
%  for (int n=0; n < nz; n+=4) {
%    float* a = &dom.col_id_t[0]+n+nz*r;
%    __m512i vect_i = _mm512_setzero_ps();
%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
%
%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
%
%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
%
%    //perform first tensor multiplication
%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
%
%    //three more times with BBBB, CCCC, and DDDD
%  }
%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
%\end{lstlisting}
%\end{center}
%\end{figure}
%

%\begin{figure*}
%\begin{verbatim}
%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
%int int_mask = 0x1111;
%__mmask16 mask = _mm512_int2mask(int_mask);
%
%#pragma omp parallel for
%for (int r=0; r < nb_rows; r++) {
%  __m512 accu = _mm512_setzero_ps();
%
%  for (int n=0; n < nz; n+=4) {
%    float* a = &dom.col_id_t[0]+n+nz*r;
%    __m512i vect_i = _mm512_setzero_ps();
%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
%
%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
%
%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
%
%    //perform first tensor multiplication
%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
%
%    //three more times with BBBB, CCCC, and DDDD
%  }
%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
%} 
%\end{verbatim}
%\caption{mat mul code.}
%\label{code:mat_mul}
%\end{figure*}

>>>>

As explained in Section~\ref{sec:model}, we expect to achieve a performance of
about 210Gflop/s in single precision in the best case. We will focus
on the single precision case with 4 vectors and 4 matrices. But
similar techniques apply for other combinations. This target
performance represents 10\% of the peak performance of the
architecture which is impossible without an efficient vectorization of
the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
instruction on fully loaded registers must be executed at most every
10 cycles.

@@

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for 4 non zeroes of
the matrix at once. The 4 floats coming from each matrix four 4
consecutive non zeroes are naturally grouped within each lane of the
SIMD vector. One can use a swizzle operation to broadcast the 4
elements of a single matrix to take the whole vector (values
$A_h^1,A_h^2,A_h^3,A_h^4$ are broadcasted in Figure~\ref{code:mat_mul}).

Loading the entries from the {\tt x} vector is a little bit more
complicated. First the columns pointers of 4 different non zeroes are
loaded in a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to unpack for the
  LSB and MSB. Both are mandatory even if one is known not to be
  useful according the documentation of the hardware}. (The column pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They replicated
so that each fill a vector lane using a swizzle operation. Then each
value is multiplied by 4 (because there are 4 vectors) and the offsets
$(3,2,1,0)$ is added each lane to obtain the correct index of the
elements of the {\tt x} vector within each lane. Then a {\tt gather}
operation is performed bring the 16 entries of the {\tt x} vector into
the SIMD register. At this point we have a SIMD register where each
lane is the four vector entries for a non zero element. Using a
broadcast permutation, one can replicate an entry of the four vectors
accross each lane.
